---
title: "Getting started with satpt"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with satpt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

With the package installed and loaded into the current *R* environment, we may begin exploring the functionality of this package through examples.

```{r}
library(satpt)
```

# Examples

## Basic usage of `satpt`
To start, we will load in the example [*Difficult Diagnoses*](https://deboonstra.github.io/satpt/reference/diagnoses.html) data to work with.

```{r}
# Load diagnoses data
data(diagnoses)
str(diagnoses)

## Turn diagnoses data into data.frame
diagnoses <- do.call(what = "cbind.data.frame", args = diagnoses)
```

We will start by examining whether the responses collected in the second question (`q2`) of the *diagnoses* data achieved saturation. 

```{r}
levels(diagnoses$q2)
```

The `res` object below is a `satpt` object that contains the results of the saturation point analysis and is structurely a *R* `list` with twelve elements produced by main function `satpt::satpt()`. You may see the [documentation of `satpt::satpt()`](https://deboonstra.github.io/satpt/reference/satpt.html) for a detailed description of all the elements of a `satpt` object.

```{r}
res <- satpt(y = diagnoses$q2)
```

For a basic examination of the saturation point analysis, you may simply print the results object. This will provide information of whether the maximum standard errors of the response categories are less than or equal to the saturation threshold along with sample proportions and standard errors of the response categories. In this example, the default saturation threshold (about 0.0255 — see *Saturation thresholds* below) was used and the maximum standard error was less than that threshold. Thus, saturation of the response categories is achieved.

```{r}
print(res)
```

When a more detailed look at the analysis is warranted, simply call [`summary()`](https://deboonstra.github.io/satpt/reference/satpt.html) on the returned `satpt` object, as seen below.

```{r}
summary(res)
```

The goal of [`summary()`](https://deboonstra.github.io/satpt/reference/satpt.html) is to provide a conditional comphrensive overview of the saturation point analysis similar in style to results produced by *SAS* procedures.

Another way of examining saturation point analysis is graphical by plotting the standard errors of the overall sample proportions for each response category in relation to the saturation threshold. With all the standard errors falling below the saturation threshold, we would say that we have achieved saturation for these responses.

```{r}
graphics::par(oma = c(0, 0, 0, 8))
plot(res)
# adding legend
legend_right(
  legend = "Saturation\nthreshold",
  col = "firebrick", lty = 3, lwd = 2,
  cex = 0.75
)
```

## Possibility of response bias

There will be times the data collection process leads to responses with the potential of having response bias. Thus, when performing saturation point analysis we must test for response bias and if response bias is present, we must account for that bias in the saturation point analysis. One example where response could be present is collecting data in waves based on non-responses in the previous waves. The *diagnoses* data previously explored was collected in this manner. Presented below is the number of responses collected for the second question for each data collection `wave`.

```{r}
stats::ftable(x = diagnoses$wave)
```

So, we are going to examine the responses collected in each wave for saturation, where the previous wave of responses will be included in the current wave and the package's default *saturation threshold* is used.

### First data collection period

During the first data collection period, we are going to assume response bias is not possible because the responses were **not** collected due to non-responses in the previous wave, as there was no "previous" data collected.

```{r}
# Subsetting the data to only include data collected during the first wave
diagnoses1 <- subset(x = diagnoses, subset = wave == 1)

# Performing saturation point analysis
res1 <- satpt(
  y = diagnoses1$q2,
  dimnames = "Responses to survey"
)
print(res1)
```

### Second data collection period

The results from the first data collection period show that saturation was not achieved, as the maximum standard error of the response items was not less than or equal to the saturation threshold. So, we'll examine the data collected during the first and second data collection period for saturation of responses. For this analysis, we must account for the possibility of response bias by specifying the `by` argument in [`satpt::satpt()`](https://deboonstra.github.io/satpt/reference/satpt.html) because the responses collected during the second wave include responses from *EIN* members who did not respond previously.

```{r}
# Subsetting the data to only include data collected during the first and second
# waves of data
diagnoses2 <- subset(
  x = diagnoses,
  subset = wave %in% c(1, 2)
)

# Performing saturation point analysis
res2 <- satpt(
  y = diagnoses2$q2,
  by = diagnoses2$wave,
  dimnames = c("by" = "Data collection period", "y" = "Responses to survey")
)
print(res2)
```

With the inclusion of the responses from the second data collection, saturation is achieved. We can take a look at the summary of the analysis, including the test of independence of the responses.

```{r}
summary(res2)
```

The above Pearson's $\chi^{2}$ test for independence shows that responses bias was not present with a non-significant p-value. Thus, signifying the responses collected during the first and second data collection waves are not statistically different. With no response bias being present, traditional standard errors for the overall sample proportions of the response categories were calculated.

# Saturation thresholds

The default saturation threshold in [`satpt::satpt()`](https://deboonstra.github.io/satpt/reference/satpt.html) is `0.05 / qnorm(0.975)` (about `0.0255`) because this is the standard error that produces a $95\%$ confidence interval of total width 0.1 around each sample proportion. More specifically, for a $95\%$ confidence interval with a margin of error (**ME**) of 0.05,

$$
ME = q \times \text{SE} = 0.05 \quad \Rightarrow \quad \text{SE} = \frac{0.05}{q}
$$

where $q = z_{0.975} \approx 1.96$ is the relevant quantile of the standard normal. Solving for SE gives the saturation threshold $\approx 0.0255$. The table below provides more *saturation thresholds* based on confidence intervals that have a total width of 0.1.

```{r, echo = FALSE}
if (requireNamespace("kableExtra", quietly = TRUE)) {
  library(kableExtra)
  ci_levels <- c(0.90, 0.95, 0.99)
  dat <- data.frame(
    ci_level = paste0(format(100 * ci_levels), "%"),
    threshold = 0.05 / qnorm(p = (1 + ci_levels) / 2)
  )
  tab <- knitr::kable(
    x = dat,
    format = "html",
    digits = 4,
    col.names = c("Confidence level", "Saturation threshold"),
    align = "cc"
  )
  kableExtra::kable_styling(
    kable_input = tab,
    bootstrap_options = c("hover"),
    full_width = FALSE,
    position = "center"
  )
}
```

The saturation threshold may be as small or as large as one wants. However, the threshold must never exceed 0.25 or be lower than 0 because the maximum standard error for any sample proportion is 0.25 when the sample proportion is 0.5.
