---
title: "Select All that Apply Questions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Select All that Apply Questions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Not every survey question will contain responses that are mutually exclusive. Specifically, it is quite common for surveys to include questions that state, "Select all the responses that all." An example of this type of question can be seen on the first question of the *EIN* survey titled [*Difficult Diagnoses*](https://ein.idsociety.org/surveys/survey/166/). So, the focus of this document will be explore we can use `satpt` to determine if saturation has been achieved for these types of questions. 

We'll start off by call the `library` function to `satpt` to obtain all the functionality of the package

```{r}
library(satpt)
```

and load in the example [*Difficult Diagnoses*](https://deboonstra.github.io/satpt/reference/diagnoses.html) data.

```{r}
data(diagnoses)
```

When we examine the first six and last six of the responses to question one, we see the responses to this "select that all apply" question are separated by the vertical pipe operator (`"|"`). 

```{r}
head(diagnoses$q1)
tail(diagnoses$q1)
```

From the original survey, [*Difficult Diagnoses*](https://ein.idsociety.org/surveys/survey/166/), we know the reported values relate to survey responses in this manner.

- `"Broadrange"`: Broad range 16S rRNA gene sequencing
- `"Wholegenome"`: Whole genome sequencing
- `"MNGS"`: Metagenomic next-generation sequencing (mNGS)
- `"None"`: None of the above
- `"Other"`: Other, specify

To analyze each response item separately we need to convert this single string column into a `data.frame` with one indicator column per response. The exported [`split_select_all_apply()`](https://deboonstra.github.io/satpt/reference/split_select_all_apply.html) helper does exactly this:

```{r}
q1 <- split_select_all_apply(x = diagnoses$q1, sep = "|")
str(q1)
```

The `"select all that apply"` responses are now a collection of indicator columns, where `"Yes"` denotes that response field was selected and `"No"` denotes it was not. These responses were collected in different data collection periods due to non-response in earlier waves, so we use the `wave` variable in `diagnoses` to control for potential response bias.

Pass the indicator data.frame straight to [`satpt::satpt()`](https://deboonstra.github.io/satpt/reference/satpt.html) with `select_all_apply = TRUE`:

```{r}
res <- satpt(
  y = q1,
  by = diagnoses$wave,
  select_all_apply = TRUE,
  dimnames = c(
    "by" = "Data collection period",
    "y" = "Responses to survey"
  )
)
```

Each response field from the "select all that apply" question achieved saturation with the default saturation threshold (about 0.0255, the standard error that produces a 95% CI of width 0.1), so the question has reached saturation.

```{r}
summary(res)
```

If you are running a whole-survey analysis with [`satpt::satpt_survey()`](https://deboonstra.github.io/satpt/reference/satpt_survey.html), pass the same separator inline via the `split` argument and the conversion happens automatically:

```{r, eval = FALSE}
satpt_survey(diagnoses, by = "wave", split = list(q1 = "|"))
```
